\addcontentsline{toc}{chapter}{Publications covered as part of thesis}
\chapter*{Publications covered as part of thesis}
\begin{enumerate}
    \item Khurana, V., Kumar, Y., Hollenstein, N., Kumar, R., \& Krishnamurthy, B. (2023). Synthesizing Human Gaze Feedback for Improved NLP Performance. In Proceedings of the 17th Conference of the European Chapter of the Association for Computational Linguistics (pp. 1895-1908).
    
    \item Kumar, Y., Jha, R., Gupta, A., Aggarwal, M., Garg, A., Malyan, T., Bhardwaj, A., Ratn Shah, R., Krishnamurthy, B., \& Chen, C. (2023). Persuasion Strategies in Advertisements. Proceedings of the AAAI Conference on Artificial Intelligence, 37(1), 57-66. \url{https://doi.org/10.1609/aaai.v37i1.25076}

    \item Bhattacharya, A., Singla, Y. K., Krishnamurthy, B., Shah, R. R., \& Chen, C. (2023). A Video Is Worth 4096 Tokens: Verbalize Videos To Understand Them In Zero Shot. In Proceedings of the 2023 Conference on Empirical Methods in Natural Language Processing, pages 9822–9839, Singapore. Association for Computational Linguistics. (\textbf{Nominated for the best paper award!})

    \item Khandelwal, A., Agrawal, A., Bhattacharyya, A., Singla, Y.K., Singh, S., Bhattacharya, U., Dasgupta, I., Petrangeli, S., Shah, R.R., Chen, C. and Krishnamurthy, B., 2024. Large Content And Behavior Models To Understand, Simulate, And Optimize Content And Behavior. International Conference on Learning Representations. (\textbf{Spotlight and nominated for award!})

    \item  S I, H., Singh, S., K Singla, Y., Krishnamurthy, B., Chen, C., Baths V., and Ratn Shah, R. (2024). Long-Term Ad Memorability: Understanding and Generating Memorable Ads. Proceedings of the Winter Conference on Applications of Computer Vision (WACV).

    \item Khurana, V., Singla, Y.K., Subramanian, J., Shah, R.R., Chen, C., Xu, Z. and Krishnamurthy, B., 2023. Measuring and Improving Engagement of Text-to-Image Generation Models. arXiv preprint arXiv:2311.10995. (Under review)

    \item Singh, S., SI, H., Singla, Y. K., Baths, V., Shah, R. R., Chen, C., and Krishnamurthy, B. (2024). Teaching Human Behavior Improves Content Understanding Abilities Of VLMs. arXiv preprint arXiv:2405.00942. (Under review)

\end{enumerate}


\addcontentsline{toc}{chapter}{Dataset contributions covered as part of this thesis}
\chapter*{Dataset contributions covered as part of this thesis}

\begin{enumerate}
    \item Persuasion Strategies for Images: \url{https://midas-research.github.io/persuasion-advertisements/}
    \item Persuasion Strategies for Videos: \url{https://behavior-in-the-wild.github.io/video-4096.html}
    \item Content Behavior Corpus: \url{https://behavior-in-the-wild.github.io/LCBM.html}
    \item Long-term memorability of Advertisements:
    \begin{enumerate}
        \item LAMBDA: \url{https://behavior-in-the-wild.github.io/memorability.html}
        \item UltraLAMBDA: \url{https://behavior-in-the-wild.github.io/memorability.html}
    \end{enumerate}
    \item EngagingImageNet: \url{https://behavior-in-the-wild.github.io/measure-engagement}
\end{enumerate}




\addcontentsline{toc}{chapter}{Other Publications}
\chapter*{Other Publications}

\begin{enumerate}
    \item Bhattacharyya, A., Singla, Y. K., Aggarwal, S., Menta, T., SR, N., \& Krishnamurthy, B. (2025). Align via actions: Learning behavior aligns LLMs with human opinions in zero-shot. Association for Computational Linguistics Rolling Review. \textbf{Nominated for Best Paper Award}.

    \item Choudhary, N., Goyal, P., Siwatch, D., Chandak, A., Mahajan, H., Khurana, V., \& Singla, Y. K. (2025). AdQuestA: Knowledge-guided visual question answer framework for advertisements. Proceedings of the Winter Conference on Applications of Computer Vision (WACV).

    \item Patnaik, S., Changwal, H., Aggarwal, M., Bhatia, S., Singla, Y. K., \& Krishnamurthy, B. (2024). CABINET: Content relevance-based noise reduction for table question answering. Proceedings of the International Conference on Learning Representations (ICLR). \textbf{Spotlight}.

    \item Anand, A., Nair, A., Prasad, K., Narayan, V., Lal, N., Mahata, D., Singla, Y. K., \& Shah, R. (2024). Advances in citation text generation: Leveraging multi-source Seq2Seq models and large language models. Proceedings of the Conference on Information and Knowledge Management (CIKM).

    \item Singla, Y. K. (2023). Can we use some advances in AI to teach AI? How could we make AI education more interdisciplinary? AI Matters as part of EAAI-23 Blue Sky Ideas in Artificial Intelligence Education from the AAAI/ACM SIGAI New and Future AI Educator Program.

    \item Singla, Y. K., Singh, S., Parekh, S., Li, J. J., Shah, R. R., \& Chen, C. (2023). Automatic essay scoring systems are both overstable and oversensitive: Explaining why and proposing defenses. Dialogue and Discourse Journal.

    \item S., S., Pupneja, A., Mital, S., Shah, C., Bawkar, M., Gupta, L. P., Kumar, A., Singla, Y. K., Gupta, R., \& Shah, R. R. (2023). H-AES: Towards automated essay scoring for Hindi. Proceedings of the Educational Advances in Artificial Intelligence (EAAI) at AAAI.

    \item Singla, Y. K., Parekh, S., Singh, S., Chen, C., Krishnamurthy, B., \& Shah, R. R. (2022). MINIMAL: Mining models for data-free universal adversarial triggers. Proceedings of the AAAI Conference on Artificial Intelligence (AAAI).

    \item Ghosh, S., Kumar, S., Singla, Y. K., Shah, R. R., \& Umesh, S. (2022). Span classification with structured information for disfluency detection in spoken utterances. Proceedings of Interspeech.

    \item Singla, Y. K., Krishna, S., Shah, R. R., \& Chen, C. (2022). Using sampling to estimate and improve performance of automated scoring systems with guarantees. Proceedings of the AAAI Conference on Artificial Intelligence - Educational Advances in Artificial Intelligence (AAAI-EAAI).

    \item Grover, M., Bamdev, P., Singla, Y. K., Vafaee, P., Hama, M., \& Shah, R. R. (2022). Automated speech scoring system under the lens: Evaluating and interpreting the linguistic cues for language proficiency. International Journal of Artificial Intelligence in Education.

    \item Mathur, A. N., Kumar, Y., Batra, D., Shah, R. R., Zimmermann, R., \& Chen, C. (2021). Lifi: Towards linguistically informed frame interpolation. Proceedings of the IEEE International Conference on Acoustics, Speech, and Signal Processing (ICASSP).

    \item Singla, Y. K., Gupta, A., Bagga, S., Chen, C., Krishnamurthy, B., \& Shah, R. R. (2021). Speaker-conditioned hierarchical modeling for automated speech scoring. Proceedings of the Conference on Information and Knowledge Management (CIKM).

    \item Sikka, J., Satya, K., Kumar, Y., Uppal, S., Shah, R. R., \& Zimmermann, R. (2020). Learning-based methods for code runtime complexity prediction. Proceedings of the European Conference on Information Retrieval (ECIR).

    \item Sahrawat, D., Mahata, D., Zhang, R., Kulkarni, M., Sharma, A., Gosangi, R., Stent, A., Kumar, Y., Shah, R. R., \& Zimmermann, R. (2020). Keyphrase extraction as a sequence labeling task using transformers. Proceedings of the European Conference on Information Retrieval (ECIR).

    \item Kumar, Y., Sahrawat, D., Maheshwari, S., Mahata, D., Shah, R. R., Yin, Y., Zimmermann, R., \& Stent, A. (2020). Harnessing GANs for zero-shot learning of new classes in visual speech recognition. Proceedings of the AAAI Conference on Artificial Intelligence (AAAI).

    \item Srivastava, N., Saxena, A., Kumar, Y., Mahata, D., Shah, R. R., Stent, A., \& Zimmermann, R. (2019). MobiVSR - Mobile application for visual speech recognition. Proceedings of Interspeech.

    \item Uttam, S., Kumar, Y., Sahrawat, D., Aggarwal, M., Mahata, D., Shah, R. R., \& Stent, A. (2019). Hush-Hush speak: Speech reconstruction using silent videos. Proceedings of Interspeech.

    \item Kumar, Y., Jain, R., Mohd. Salik, K., Shah, R. R., Yin, Y., \& Zimmermann, R. (2019). Lipper: Synthesizing thy speech using multi-view lipreading. Proceedings of the AAAI Conference on Artificial Intelligence (AAAI).

    \item Kumar, Y., Aggarwal, S., Mahata, D., Shah, R. R., Kumaraguru, P., \& Zimmermann, R. (2019). Get IT scored using AutoSAS - An automated system for scoring short answers. Proceedings of the Educational Advances in Artificial Intelligence (EAAI) at AAAI.

    \item Kumar, Y., Jain, R., Mohd. Salik, K., Shah, R. R., Yin, Y., \& Zimmermann, R. (2018). MyLipper: A personalized system for speech reconstruction using multi-view visual feeds. Proceedings of the International Symposium on Multimedia (ISM). \textbf{Best Poster Runner-up}.

    \item Kumar, Y., Aggarwal, M., Nawal, P., Satoh, S., Shah, R. R., \& Zimmermann, R. (2018). Harnessing AI for speech reconstruction using multi-view silent video feed. Proceedings of the ACM Multimedia Conference (ACMMM).
\end{enumerate}



\addcontentsline{toc}{chapter}{Patents}
\chapter*{Patents}

\begin{enumerate}
    \item Kumar, Y., Singh, S., Park, S., Prasoon, P., Sainath, N., Joshi, N. S., Srikanth, N., Puri, N., Aggarwal, M., Subramanian, J., Palwe, G., Krishnamurthy, B., Rozen, M. W., Naware, M., \& Chung, H. (2024). Digital content analysis (U.S. Patent App. No. 20240355020). U.S. Patent and Trademark Office.

    \item Kumar, Y., \& Khurana, V. (2024). Systems and methods for generating scanpaths (U.S. Patent App. No. 18/109,990). U.S. Patent and Trademark Office.

    \item Kumar, Y., Khuc, V. N., Srivastava, V., Moorarka, U., Verma, S., Shahid, S., Bansal, S., Venkitachalam, S., Steimer, S., Karmakar, S., Srivastav, N., Puri, N., Naware, M., Jain, K. K., Singh, K. M., Chung, H., Bacila, H., Lordache, F. S., Pai, D., \& Krishnamurthy, B. (2024). Determining user affinities for content generation applications (U.S. Patent No. 12,008,033). U.S. Patent and Trademark Office.

    \item Kumar, Y., Ahlawat, V., Zhang, R., Aggarwal, M., Palwe, G. K., Krishnamurthy, B., \& Khurana, V. (2024). Attention aware multi-modal model for content understanding (U.S. Patent App. No. 17/944,502). U.S. Patent and Trademark Office.
    
    \item Kumar, Y., Singh, S., George, W. B., Liu, T. C., Basetty, S., Prasoon, P., Puri, N., Naware, M., Corlan, M., Butikofer, J. M., Chauhan, A., Singh, K. M., O'Reilly, J. P., Chung, H., Dest, L., Goudie-Nice, C. H., Pack, B. J., Krishnamurthy, B., Jain, K. K., Klimetschek, A., \& Rozen, M. W. (2024). Content analytics as part of content creation (U.S. Patent No. 11,907,508 B1). U.S. Patent and Trademark Office.
    
    \item Kumar, Y., \& Krishnamurthy, B. (2023). Visual speech recognition for digital videos utilizing generative adversarial learning (U.S. Patent App. No. 17/650,020; CN Patent App. No. 202211407981.2A; DE Patent App. No. 102022131824.9A). U.S. Patent and Trademark Office, China National Intellectual Property Administration, \& German Patent and Trade Mark Office.

    \item Kumar, Y. (2021). Pose-invariant visual speech recognition using a single view input (U.S. Patent No. 10,937,428). U.S. Patent and Trademark Office.

\end{enumerate}


\addcontentsline{toc}{chapter}{Awards}
\chapter*{Awards}

\begin{itemize}
     \item{Google PhD Fellow (1 among 40 candidates all over the world)}
    \item{The Indian Prime Minister PhD Fellowship (1 among 38 candidates all over India across all STEAM branches)}
    \item{University at Buffalo - PhD Best First-Year Achiever Award}
    \item{IIIT-Delhi Dean IRD Research Excellence Award for my PhD research for three consecutive years (2021-23)}
   \item{Adobe Outstanding Young Engineer Award}
    \item{Two spotlight paper awards at ICLR-24}
    \item{Nominated for Best Paper Award at EMNLP-23}
    \item{Best paper award nomination ACL Rolling Review-24}
    \item {Best poster paper runner up (\textit{i.e.} 2nd position) award at ISM-2018}
    \item {Best Student Abstract Research Paper at AAAI-2019}
    \item{Best Adobe Sneaks Award - 2021 \href{https://business.adobe.com/summit/2021/sessions/catchy-content-sneak-gs3-4.html}{Project \#CatchyContent}. News Coverage: \href{https://go.forrester.com/blogs/adobe-catchy-content-shows-how-content-intelligence-powers-personalization/}{Forrester}, \href{https://www.fastcompany.com/90629922/why-is-nobody-clicking-your-website-adobes-brutally-honest-new-ai-can-tell-you}{Fast Company}, and other media and research houses}

    %\item{Awarded Overseas Research Fellowship Award by IIIT-Delhi for the year 2024}
    \item{Awarded AI Future Educator Travel Award by EAAI/AAAI-23}
    \item{Heidelberg Laureate Young Researcher - 2021}
    \item{SIGIR Student Grant-2021}
    %\item {Interspeech Young Scientists Grant-2019}
    %\item {EAAI-AAAI scholarship at AAAI-2019}
    %\item {AAAI student scholarship at AAAI-2019, AAAI-2020}
    \item {Winner (\textit{i.e.} 1st position) of ACM All-India Student Chapters Challenge 2019}
\end{itemize}
